### 2016-10-19 R/ITHIM.Rmd
Tweaked Figure 1 from methods paper
Created two new figures
Also created appendix.Rnw to store old R code.

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
# install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r","devtools"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
#library("cowplot")
#library("grid")
#library("scales")
#library("ggthemes")
```

@@

#upper <- 10

@@

#geom_hline(yintercept=seq(1, upper, 1), col="white", lwd=1) +
# + theme_tufte(base_size=12, ticks=T)
#
# + annotate("text", x = 3.5, y = 5, adj=1,  family="serif", label = c("Foo"))
# + scale_y_continuous(name="DALYs Averted Nationwide (Thousands)") + facet_grid( Disease ~ State) + coord_flip() # + theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=8))
```

@@

## New Analysis

@@

ITHIM.scenario <- ITHIM.national
ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=2)/10)
ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=2)/10)

@@

DALYs <- c(); DALYs.quints <- c();
atVec <- seq(45, 120, length.out = 5)
muNonTravelVec <- c(1,5,10,25)
DTotal <- data.frame()

@@

#D <- rbind(data.frame(AT = atVec, DALYs = DALYs, quants = "percentiles"),data.frame(AT = atVec, DALYs = DALYs.quints, quants = "quintiles"))
#+ scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + facet_grid( . ~ quints)
#p + stat_smooth( method = loess) + theme_bw() + scale_x_continuous(name="Mean Non-Travel METs (MET-hrs./week) (Referent Group)") + scale_y_continuous(name="Change in National DALYs (Thousands)") + coord_cartesian(xlim = c(0, upper)) + geom_point() + facet_grid( . ~ quints)
```

# Appendix

## Non-Travel METs

```{r newMETs, echo = TRUE, eval = FALSE, echo = FALSE}
newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
dataColumns <- c(1,5)
D <- newMETs[,dataColumns]
metaData <- newMETs[,-dataColumns]
newMETs <- list(Data = D, metaData = metaData)
activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))

rownames(newMETs$Data) <- activityCode
rownames(newMETs$metaData) <- activityCode

newMETs$metaData <- within(newMETs$metaData, {
  Travel.related..Y.N. <- as.factor(Travel.related..Y.N.)
})

metKey <- newMETs$Data$AinsworthMETS
names(metKey) <- activityCode

@@

```{r readnewMETs, echo = TRUE, eval = FALSE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat"
summary <- read.table(summaryFile, header=T, skip=0, sep=",")
metaDataColumns <- 1:24
summaryList <- list(Data = summary[,-metaDataColumns], metaData = summary[,metaDataColumns])

summaryList$metaData <- within(summaryList$metaData,{
    TESEX <- as.factor(ifelse(TESEX == 1, "male", ifelse(TESEX == 2, "female",NA)))
    AGECAT <- ifelse(TEAGE <5, "00-04",
                     ifelse(TEAGE > 4 & TEAGE < 15, "05-14",
                     ifelse(TEAGE > 14 & TEAGE < 30, "15-29",
                     ifelse(TEAGE > 29 & TEAGE < 45, "30-44",
                     ifelse(TEAGE > 44 & TEAGE < 60, "45-59",
                     ifelse(TEAGE > 59 & TEAGE < 70, "60-69",
                     ifelse(TEAGE > 69 & TEAGE < 80, "70-79", "80+"
                            )))))))})

We shift the MET estimates by the minimum value so that the new minimum MET estimate is zero.

```{r metKey, echo = TRUE, eval = FALSE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

activities <- nonTravelActivities

metKeyVector <- (metKey[activities] - min(metKey, na.rm = TRUE))/60*7
names(metKeyVector) <- activities
metKeyMatrix <- matrix(metKeyVector, nrow = nrow(summaryList$Data), ncol = length(metKeyVector), byrow = TRUE)
dimnames(metKeyMatrix) <- list( NULL, activities)

METs <- data.frame(METs = rowSums(summaryList$Data[,activities] * metKeyMatrix[,activities], na.rm = TRUE), AGE = summaryList$metaData$TEAGE, SEX = summaryList$metaData$TESEX, AGECAT = summaryList$metaData$AGECAT, row.names = summaryList$metaData$TUCASEID )

D <- merge(cps, METs, by.y = "row.names", by.x = "TUCASEID")
ggplot(D, aes(x = METs, ..density.., colour = AGECAT)) + geom_freqpoly(binwidth = 10) + facet_grid( SEX ~ .) + labs(x = "Non-Travel Marginal MET-hrs./week", colour = "Age") + coord_cartesian(xlim = c(0, 400))

```{r parameters, eval = FALSE, echo = FALSE}
parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             )

METmatrix <- rbind(matrix(0,nrow = 2, ncol = 2), matrix(parameters$mean, ncol = 2))
METmatrix <- METmatrix/METmatrix[3,2]
```

```{r roadInjuries, eval = FALSE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
convertState <- function( stateCode ){

stateCode <- as.character(stateCode)

stateCodeMatrix <- matrix(c("Alabama","1","Nebraska","31","Alaska","2","Nevada","32","Arizona","4","New Hampshire","33","Arkansas","5","New Jersey","34","California","6","New Mexico","35","Colorado","8","New York","36","Connecticut","9","North Carolina","37","Delaware","10","North Dakota","38","District of Columbia","11", "Ohio", "39","Florida","12","Oklahoma","40","Georgia","13","Oregon","41","Hawaii","15","Pennsylvania","42","Idaho","16","Puerto Rico","43","Illinois","17","Rhode Island","44","Indiana","18","South Carolina","45","Iowa","19","South Dakota","46","Kansas","20","Tennessee","47","Kentucky","21","Texas","48","Louisiana","22","Utah","49","Maine","23","Vermont","50","Maryland","24","Virgin Islands (since 2004)","52","Massachusetts","25","Virginia","51","Michigan","26","Washington","53","Minnesota","27","West Virginia","54","Mississippi","28","Wisconsin","55","Missouri","29","Wyoming","56","Montana","30"), byrow = TRUE, ncol = 2)

stateAbbrevMatrix <- matrix(c("Alabama","AL","Alaska","AK","Arizona","AZ","Arkansas","AR","California","CA","Colorado","CO","Connecticut","CT","Delaware","DE","District of Columbia","DC","Florida","FL","Georgia","GA","Hawaii","HI","Idaho","ID","Illinois","IL","Indiana","IN","Iowa","IA","Kansas","KS","Kentucky","KY","Louisiana","LA","Maine","ME","Maryland","MD","Massachusetts","MA","Michigan","MI","Minnesota","MN","Mississippi","MS","Missouri","MO","Montana","MT","Nebraska","NE","Nevada","NV","New Hampshire","NH","New Jersey","NJ","New Mexico","NM","New York","NY","North Carolina","NC"," North Dakota","ND","Ohio","OH","Oklahoma","OK","Oregon","OR","Pennsylvania","PA","Rhode Island","RI","South Carolina","SC","South Dakota","SD","Tennessee","TN","Texas","TX","Utah","UT","Vermont","VT","Virginia","VA","Washington","WA","West Virginia","WV","Wisconsin","WI","Wyoming","WY"), byrow = TRUE, ncol = 2)

stateNames <- stateCodeMatrix[,1]
names(stateNames) <- stateCodeMatrix[,2]

stateAbbrevs <- stateAbbrevMatrix[,2]
names(stateAbbrevs) <- stateAbbrevMatrix[,1]

return(stateAbbrevs[stateNames[stateCode]])

}

roadFatalaties <- read.csv(file = "./../data/FARS/BikePedDeaths99-13.csv", na.strings = c("998","999"))
roadFatalaties <- within(roadFatalaties, {
  ageClass <- convertToAgeClass(AGE)
  SEX <- ifelse(SEX == 1, "M", ifelse(SEX == 2, "F", NA))
  SEX <- factor(SEX, levels = c("F","M"))
  STATE <- as.factor(convertState(STATE))
})
roadFatalaties <- subset(roadFatalaties, !is.na(SEX) ) #& STATE %in% c("California", "New York"))

p <- ggplot(roadFatalaties, aes(Type, fill = STATE))
p + geom_bar(position = "dodge") + theme_bw()+ theme(axis.text.x  = element_text(angle=90, vjust=0.5))# + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) #+ facet_grid( Type ~ ., scales = "free_y")

walkTime <- unlist(lapply(meanWalkList, function(x) x[3,2]))

foo <- data.frame(with(roadFatalaties, table(STATE,Type)))
within(subset(foo, Type == "Ped"), Walk <- walkTime[as.character(STATE)])

### CDC Mortality

Next we parse the output of a CDC Wonder query to estimate mortality rates.

The data files in ~/Box Sync/work/CoBenefits/MaggieShare/VitalStats/CDCwonder/.

#### Query Results File (HHSmetro_Age_sex_2010-14.csv)

```{r CDC, eval = FALSE, echo = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

wonder <- subset(wonder, WONDERageCode != "NS") # remove 'Not stated' Age entries
```

The following two recoding chunks should probably be redone.  They
depend on the ordering of the factors which seems dangerous.

```{r ITHIMCause, eval = FALSE, echo = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = FALSE, echo = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

Depression was not in the WONDER data, since there are no deaths, have to add it manually.

```{r wonder, eval = FALSE, echo = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

Replace suppressed values with random number 1-9

```{r suppressed, eval = FALSE, echo = FALSE, warning = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

10 HHS x 2 sex x 8 age x 13 disease = 2,080 strata.

```{r summary2, eval = FALSE, echo = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

#### Single Year Data File (HHSmetro_Age_sex_2014_AllCause.csv)

Read in data for single year population.  Processed in Excel, to
combine two WONDER queries, need to go back and script the processing
of the raw WONDER output

```{r singleyear, eval = FALSE, echo = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

names(pop14) <- c("HHSRegion", "HHSRegionCode","gender", "genderCode","WONDERage","WONDERageCode", "deaths2014AllCause", "population2014", "crudeRate2014","random1","random2")
```
This is delicate(read sloppy), really need to check the order of age groups returned by the 'unique(pop14$WONDERage)' command.

```{r sloppy, eval = FALSE, echo = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = FALSE, echo = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

popSummary <- ddply(pop14, .(HHSRegionCode, gender, ITHIMage), summarise,
                    population2014 = sum(as.numeric(as.character(population2014)), na.rm=T),
                   AllCauseDeaths2014 = sum(as.numeric(as.character(deaths2014AllCause)), na.rm=T)
)

HHSmetroCauseSpecific <- merge(wonderSummary, popSummary)
```
#### Read USGBD file (USGBD.csv)

Presumably this comes from an EXCEL file.

```{r USGBD, eval = FALSE, echo = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

burden <- merge(HHSmetroCauseSpecific, USGBD, all.x=T)
burden$deaths <- burden$numberOfDeaths5yr/5
## neil uses 10, and if deaths<10 sets ratio to 1
burden$ratio <- ifelse(burden$GBDdeaths == 0, 1, ((burden$deaths/burden$population2014)/(burden$GBDdeaths/burden$GBDpop)))
burden$YLL <- burden$ratio * (burden$GBDyll/burden$GBDpop) * burden$population2014
burden$YLD <- burden$ratio * (burden$GBDyld/burden$GBDpop) * burden$population2014
burden$DALY <- burden$ratio * (burden$GBDdaly/burden$GBDpop) * burden$population2014

burden$sex <- factor(ifelse(burden$gender == "Male","M","F"))
levels(burden$sex)  <- c("M","F")

burdenOut <- burden[,c("HHSRegionCode","ITHIMcause","sex","ITHIMage","deaths","YLL","YLD","DALY")]
names(burdenOut) <- c("HHS","disease","sex","ageClass","dproj","yll","yld","daly")

burdenOut <- burdenOut[order(burdenOut$HHS, burdenOut$disease, burdenOut$sex, burdenOut$ageClass),]
burdenOut <- burdenOut[,c("HHS", "disease","sex","dproj","yll","yld","daly")]
```

Now we create an R list of GBD objects named by HHS region.

```{r burdenList, eval = FALSE, echo = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

## Example (National)

```{r instantiate2, eval = FALSE, echo = FALSE}
ITHIM.baseline <- createITHIM()
ITHIM.scenario <- createITHIM()
```

```{r RRs2, eval = FALSE, echo = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show222, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show22, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show23, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show24, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

## Example

```{r example, eval = FALSE, echo = FALSE}
example("ITHIM-package")
```

### Creating Different Visions

```{r vision, eval = FALSE, echo = FALSE}
ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
```
#### YLL
```{r show233, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

## Interrogate Parameters

```{r params, eval = FALSE, echo = FALSE}
names(ITHIM.baseline$parameters)
ITHIM.baseline$parameters$NonTravelMETs
ITHIM.baseline$parameters$Rwt
```

## Appendix

### DETAILS ON WONDER QUERY
Dataset: Underlying Cause of Death, 1999-2014
Query Parameters:
Title: ITHIM2014_HHS_urb_age_sex_16years
2013 Urbanization: All
Autopsy: All
Five-Year Age Groups: All
Gender: Female, Male
HHS Regions: All
Hispanic Origin: All
ICD-10 113 Cause List: Malignant neoplasms of colon, rectum and anus (C18-C21), Malignant neoplasms of trachea, bronchus and
lung (C33-C34), Malignant neoplasm of breast (C50), #Diabetes mellitus (E10-E14), #Alzheimer's disease (G30), Other forms of
chronic ischemic heart disease (I20,I25), Acute and subacute endocarditis (I33), Diseases of pericardium and acute myocarditis
(I30-I31,I40), #Essential hypertension and hypertensive renal disease
(I10,I12,I15), #Cerebrovascular diseases (I60-I69), Influenza and pneumonia (J09-J18), Other acute lower respiratory infections (J20-J22,U04), #Chronic lower respiratory diseases
(J40-J47), Other diseases of respiratory system (J00-J06,J30- J39,J67,J70-J98), Motor vehicle accidents
(V02-V04,V09.0,V09.2,V12-V14,V19.0-V19.2,V19.4-V19.6,V20-V79,V80.3-V80.5,V81.0-V81.1,V82.0-V82.1,V83-V86,V87.0-V87.8,V88.0-V88.8
 V89.0,V89.2)
Other land transport accidents
(V01,V05-V06,V09.1,V09.3-V09.9,V10-V11,V15-V18,V19.3,V19.8-V19.9,V80.0-V80.2,V80.6-V80.9,V81.2-V81.9,V82.2-V82.9,V87.9,V88.9
 V89.1,V89.3,V89.9)
Place of Death: All
Race: All
Weekday: All
Year/Month: All
Group By: Five-Year Age Groups, Gender, HHS Region, 2013 Urbanization, ICD-10 113 Cause List
Show Totals: Disabled
Show Zero Values: True
Show Suppressed: True
Calculate Rates Per: 100,000
Rate Options: Default intercensal populations for years 2001-2009 (except Infant Age Groups)
---
  Help: See http://wonder.cdc.gov/wonder/help/ucd.html for more information.
---
  Query Date: Jul 19, 2016 4:46:10 PM
---
  Suggested Citation: Centers for Disease Control and Prevention, National Center for Health Statistics. Underlying Cause of Death
1999-2014 on CDC WONDER Online Database, released 2015. Data are from the Multiple Cause of Death Files, 1999-2014, as compiled
from data provided by the 57 vital statistics jurisdictions through the Vital Statistics Cooperative Program. Accessed at
http://wonder.cdc.gov/ucd-icd10.html on Jul 19, 2016 4:46:10 PM
---
  Messages:
  1. Totals and Percent of Total are disabled when data are grouped by 113 or 130 Cause Lists. Check Caveats below for more
information.
2. Selections made to any variable in [HHS Regions, 2013 Urbanization] cause equivalent selections in the other variables. More
Information: http://wonder.cdc.gov/wonder/help/faq.html#related-variables.
3. Totals are not available for these results due to suppression constraints. More Information:
  http://wonder.cdc.gov/wonder/help/faq.html#Privacy.
---
  Footnotes:
  1. A '#' symbol preceding the label indicates a rankable cause of death. More information.
---
  Caveats:
  1. Totals and Percent of Total are disabled when data are grouped by a 113 or 130 Cause List because both aggregate and detailed
values are displayed in the table. Also be aware that charts and maps containing both aggregate and detail data could be
misleading.
2. Population and rates are labeled 'Not Applicable' when they include a subset of ages 85-100+ because populations are not
available for those ages. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Ages 85-100.
3. Circumstances in Georgia for the years 2008 and 2009 have resulted in unusually high death counts for the ICD-10 cause of
death code R99, "Other ill-defined and unspecified causes of mortality." Caution should be used in interpreting these data.
More information: http://wonder.cdc.gov/wonder/help/ucd.html#Georgia-Reporting-Anomalies.
4. Circumstances in New Jersey for the year 2009 have resulted in unusually high death counts for the ICD-10 cause of death code
R99, "Other ill-defined and unspecified causes of mortality" and therefore unusually low death counts in other ICD-10 codes,
most notably R95, "Sudden Infant Death Syndrome" and X40-X49, "Unintentional poisoning." Caution should be used in
interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#New-Jersey-Reporting-Anomalies.
5. Circumstances in California resulted in unusually high death counts for the ICD-10 cause of death code R99, "Other
ill-defined and unspecified causes of mortality" for deaths occurring in years 2000 and 2001. Caution should be used in
interpreting these data. More information: http://wonder.cdc.gov/wonder/help/ucd.html#California-Reporting-Anomalies.
6. Data are Suppressed when the data meet the criteria for confidentiality constraints. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Assurance of Confidentiality.
7. Death rates are flagged as Unreliable when the rate is calculated with a numerator of 20 or less. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Unreliable.
8. Deaths of persons with Age "Not Stated" are included in "All" counts and rates, but are not distributed among age groups,
so are not included in age-specific counts, age-specific rates or in any age-adjusted rates. More information:
  http://wonder.cdc.gov/wonder/help/ucd.html#Not Stated.
9. The population figures for year 2014 are bridged-race estimates of the July 1 resident population, from the Vintage 2014
postcensal series released by NCHS on June 30, 2015. The population figures for year 2013 are bridged-race estimates of the July
1 resident population, from the Vintage 2013 postcensal series released by NCHS on June 26, 2014. The population figures for
year 2012 are bridged-race estimates of the July 1 resident population, from the Vintage 2012 postcensal series released by NCHS
on June 13, 2013. The population figures for year 2011 are bridged-race estimates of the July 1 resident population, from the
Vintage 2011 postcensal series released by NCHS on July 18, 2012. Population figures for 2010 are April 1 Census counts. The
population figures for years 2001 - 2009 are bridged-race estimates of the July 1 resident population, from the revised
intercensal county-level 2000 - 2009 series released by NCHS on October 26, 2012. Population figures for 2000 sssssare April 1 Census
counts. Population figures for 1999 are from the 1990-1999 intercensal series of July 1 estimates. Population figures for the
infant age groups are the number of live births. <br/><b>Note:</b> Rates and population figures for years 2001 - 2009 differ
slightly from previously published reports, due to use of the population estimates which were available at the time of release.
10. The population figures used in the calculation of death rates for the age group 'under 1 year' are the estimates of the
resident population that is under one year of age. More information: http://wonder.cdc.gov/wonder/help/ucd.html#Age Group.

```{r appendix}
date()
sessionInfo()
```

<!--

# Code Graveyard

### Total Distribution

```{r total, eval = FALSE, echo = FALSE}
foo <- mapply(getTotalDistribution,
  muTravel = matrix(35,nrow = 8, ncol = 2),
  cvTravel = 1.72,
  muNonTravel = matrix(100,nrow = 8, ncol = 2),
  cvNonTravel = 1,
  p = 1,
  pWalk = rep(54.4/64.2,2),
  vWalk = rep(2.88,2),
  size = 1e5, SIMPLIFY = FALSE)
```

```{r vargo, echo = FALSE, eval = FALSE}
library(ggplot2)
library(reshape2)
library(plyr)

summaryFile <- "~/GHI/data/ATUS/atussum_2003.dat" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atussum_2003.dat

cpsFile <- "~/GHI/data/ATUS/atuscps_2003.dat" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atuscps_2003.dat

stateResultFile <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults.csv

stateResultFile2 <- "~/GHI/data/stateresults.csv" # ~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/stateresults_quintile.csv"

regionResultFile <- "~/GHI/data//regionresults_quintile.csv" #~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/regionresults_quintile.csv

#readLines("~/Box Documents/work/CoBenefits/MaggieShare/AmerTimeUse/atusact_2003.dat", n=10)
# activity <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusact_2003.dat", header=T, skip=0, sep=",")

summary <- read.table(summaryFile, header=T, skip=0, sep=",")

# respondants <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusresp_2003.dat", header=T, skip=0, sep=",")

# roster <- read.table("~/Box Documents/work/CoBenefits/MaggieShare/MET Analysis/AmerTimeUse/atusrost_2003.dat", header=T, skip=0, sep=",")

CPS <- read.table(cpsFile, header=T, skip=0, sep=",")

CPS.2 <- CPS[,c("TUCASEID", "GESTFIPS", "GEREG","GEMETSTA")]
CPS.2$IDCHAR <- as.character(CPS.2[,"TUCASEID"])

summary.2 <- summary[, c("TUCASEID", "TEAGE",    "TESEX",    "t050203",  "t130101",  "t130102",  "t130103",  "t130104",  "t130106",  "t130108", "t130109", "t130111", "t130113",  "t130115", "t130116",  "t130117", "t130119",  "t130120",  "t130122",  "t130123", "t130124",  "t130126",  "t130127", "t130128", "t130130",  "t130131",  "t130132", "t130133",  "t130134", "t130135", "t130136" , "t130199" , "t139999")]
summary.2$IDCHAR <- as.character(summary.2[,"TUCASEID"])

summary.2a <- summary.2[,c(1:3, length(summary.2))]
summary.2b <- summary.2[,c(5:length(summary.2)-1)]
minutes <- rowSums(summary.2b)
summary.2a$minutes <- minutes

metkey <- as.matrix(c(3.0,7.3,5.0,6.5,7.5,3.5,5.8,7.8,6.0,8.0,3.8,7.8,8.0,5.3,7.3,9.8,6.3,7.0,7.0,5.0,6.8,3.0,3.5,5.5,6.0,4.0,6.0,3.0,3.0,3.0), nrow = 1, ncol=30)

summary.2c <- mapply(`*`,summary.2b,metkey)
mets <- rowSums(summary.2c)
summary.2a$mets <- mets

summary.2a$agecat <- ifelse(summary.2a$TEAGE <5, 1,
                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, 2,
                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, 3,
                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, 4,
                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, 5,
                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, 6,
                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, 7, 8
)))))))

summary.2a$agecat2 <- ifelse(summary.2a$TEAGE <5, "age0-4",
                     ifelse(summary.2a$TEAGE > 4 & summary.2a$TEAGE < 15, "age5-14",
                     ifelse(summary.2a$TEAGE > 14 & summary.2a$TEAGE < 30, "age15-29",
                     ifelse(summary.2a$TEAGE > 29 & summary.2a$TEAGE < 45, "age30-44",
                     ifelse(summary.2a$TEAGE > 44 & summary.2a$TEAGE < 60, "age45-59",
                     ifelse(summary.2a$TEAGE > 59 & summary.2a$TEAGE < 70, "age60-69",
                     ifelse(summary.2a$TEAGE > 69 & summary.2a$TEAGE < 80, "age70-79", "age80+"
)))))))

use <- merge(summary.2a, CPS.2, by.x="TUCASEID", by.y="TUCASEID", all.x=T, sorted=F)

use$sexcat <- ifelse(use$TESEX == 1, "male", "female")

use$regcat <- ifelse(use$GEREG == 1, "northeast",
              ifelse(use$GEREG == 2, "midwest",
              ifelse(use$GEREG == 3, "south",  "west")))

use$metcat <- ifelse(use$GEMETSTA == 1, "metro",
              ifelse(use$GEMETSTA == 2, "non-metro", "unclassified"))

use <- use[,c("minutes","mets","agecat2","GESTFIPS","sexcat","regcat","metcat")]

#head(use)

use.region <- ddply(use, .(sexcat, agecat2, regcat, metcat), summarise,
             mean_min = mean(minutes, na.rm=T),
             mean_mets = mean(mets, na.rm=T),
             n = length(minutes)
)

>>>>

```{r libs, eval = TRUE, echo = FALSE, results = "hide", warning = FALSE, error = FALSE, message = FALSE}
options(repos='http://cran.rstudio.com/')
# install.packages(c("knitr","ggplot2","dplyr","openssl","httr", "git2r","devtools"))
library("ggplot2")
library("devtools")
library("reshape2")
library("plyr")
library("dplyr")
```

@@

@@

@@

## Figure 1

@@

ITHIM.scenario <- ITHIM.national
ITHIM.national.quints <- updateITHIM(ITHIM.national, "quantiles", seq(1,9,by=1)/10)
ITHIM.scenario.quints <- updateITHIM(ITHIM.scenario, "quantiles", seq(1,9,by=1)/10)

@@

DALYs <- c(); DALYs.quints <- c();
atVec <- seq(45, 120, length.out = 5)
muNonTravelVec <- c(1, 10, 50, 100, 150, 200)
DTotal <- data.frame()

@@

@@

## Figure 2

```{r displayMeans, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- rbind(getWalkTime(ITHIM.default),getWalkSpeed(ITHIM.default),getCycleTime(ITHIM.default),getNonTravelMETs(ITHIM.default))
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(D, aes(x = ageClass, y = mu, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="Travel Time per Week (minutes)") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ ., scales = "free_y")
```

```{r displayGBD, eval = TRUE, dev=c('pdf', 'png'), fig.width = fig.width, fig.asp = 1/phi, warning = FALSE, message = FALSE, echo = FALSE}
D <- melt(ITHIM.default$parameters$GBD)
names(D) <- c("disease","sex","ageClass","type","burden", "sex2", "disease2")
D <- within(D, sex <- factor(sex, levels = c("F","M")))

p <- ggplot(subset(D,disease %in% c("BreastCancer","ColonCancer","Dementia","Depression","Diabetes")), aes(x = ageClass, y = burden, fill = sex)) + geom_bar(stat= "identity", position = "dodge")
p + theme_bw() + scale_x_discrete(name="", labels = c("0-4","5-14","15-29","30-44","45-59","60-69","70-79","80+")) + scale_y_continuous(name="") + theme(axis.text.x  = element_text(angle = 90, vjust = 0.5, size = 10)) + facet_grid( type ~ disease, scales = "free_y")
```
